---
title: "Plots for AUC statistics"
author: "Adrian Barnett"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, comment='', dpi=400)
library(dplyr)
library(tidyr)
library(stringr)
library(splines) # for smooth spline
library(ciTools) # for confidence intervals for glm
# graphics
library(ggplot2)
library(gridExtra)
g.theme = theme_bw() + theme(panel.grid.minor = element_blank())
type_colours = c("burlywood2", "dodgerblue", "goldenrod4", "firebrick2")
source('99_functions.R')

# get the data
load('data/analysis_ready.RData') # from 4_combine_results.R

# remove AUC of exactly 1 because of accuracy issues; go to 0.99 because last interval is (0.99, 1]
results = filter(results, auc <= 0.99)

# breaks for histograms
breaks = seq(0, 1.00, 0.01) # 
# which AUC bars to colour
colour = c(0.7, 0.8, 0.9)
# x-axis for AUC
x.breaks = c(0, seq(0.5,1,0.1))
minor_breaks = seq(0.05,0.95,0.1)
# lower AUC limit for plots
lower_auc_plot = 0.48
```

## Types

Table of AUC types, simplified to means or intervals.

```{r}
library(janitor)
tab = mutate(results, type = as.character(type)) %>%
  filter(type != 'Difference') %>%
  mutate(type = ifelse(type=='Pair', 'Mean', type),
         type = ifelse(type=='Lower', 'CI', type),
         type = ifelse(type=='Upper', 'CI', type)) %>%
  tabyl(type) %>%
  adorn_totals()
flextable(tab) %>%
  theme_box() %>%
  autofit()
```

## Dates

```{r}
stats = mutate(results, year = as.numeric(format(date, '%Y'))) %>%
  summarise(median = median(year),
            q1 = quantile(year, 0.25),
            q3 = quantile(year, 0.75))
flextable(stats) %>%
  theme_box() %>%
  autofit()
```

Summary statistics on included statistics as years.


## Histogram

The plot below shows a histogram of all AUC results that were a "mean" or "pair" (so the confidence interval limits are not shown). We restricted the results to those with two or more digits after the decimal place to avoid spikes appearing due to rounding.

The results are binned using a (lower, upper] limit. For example, (0.80, 0.81] is $>$ 0.80 and $\leq$ 0.81.

```{r}
### create the data for the rounded histogram
uncounted = filter(results, 
              !(type %in% c('Difference','Lower','Upper'))) 
one_digit = filter(uncounted, 
                   auc%in% seq(0,0.9,0.1),
                   digits == 1) %>% # count use of single digits
  nrow()
one_digit_percent = 100*one_digit/nrow(uncounted)
uncounted = filter(uncounted, digits > 1) %>% # avoid spikes due to rounding
  mutate(binned = my_bin(auc, digits = 2)) # from lower limit up to upper limit: (lower,upper]
for_histo = group_by(uncounted, binned) %>%
  tally() %>%
  ungroup() %>%
  mutate(fill = ifelse(binned %in% colour, 1, 2))

# smooth Poisson model
df = 4 # degrees of freedom for smooth
model = glm(n ~ ns(binned, df = df), data = for_histo, family=poisson())
ci = add_ci(for_histo, model, names = c("lower", "upper"), alpha = 0.05) %>%
  mutate(group = 1, # to avoid separate dotted lines by fill
         residual = n - pred) %>% 
  arrange(binned)

# histogram
rounded = ggplot(data = ci, aes(x = binned, y = n, fill = factor(fill), group = group))+
  geom_bar(stat='identity', col='grey22', size=0.1)+ # size for thinner lines around bars
  geom_line(data = ci, aes(x=binned, y=pred, group = group), lty=2)+ # smoothed
  #geom_line(data = ci, aes(x=binned, y=lower, group = group), lty=2)+ # CI for smooth
  #geom_line(data = ci, aes(x=binned, y=upper, group = group), lty=2)+ # CI for smooth
  scale_fill_manual(NULL, values=c('dark red','pink','red'))+
  scale_x_continuous(breaks = x.breaks + (0.01)/2, # nudge breaks to top of bar
                     labels = x.breaks,
                     expand=c(0.01,0))+
  theme_bw()+ # keep minor grid lines
  scale_y_continuous(expand=c(0, 150))+ # reduce space to x-axis
  xlab('AUC')+
  ylab('Frequency')+
  coord_cartesian(xlim=c(lower_auc_plot, NA))+ # focus on near 0.5 and above
  theme(legend.position = 'none')
rounded
# export
jpeg('figures/histogram.jpg', width=5, height=4, units='in', res=500)
print(rounded)
invisible(dev.off())
```

The dotted line shows a smooth natural spline with `r df` degrees of freedom fitted to the counts in each bin.

There were `r one_digit` (`r round(one_digit_percent, 1)`) values excluded because they were given to just 1 decimal place.

#### Residual plot

The plot below is the residuals from the smooth fit to the distribution.

```{r}
residual = ggplot(data = ci, aes(x = binned, y = residual, fill = factor(fill)))+
  geom_bar(stat='identity', col='grey22', size=0.1)+
  scale_fill_manual(NULL, values=c('dark red','pink','red'))+
  scale_x_continuous(breaks = x.breaks + (0.01)/2, # nudge breaks to top of bar
                     labels = x.breaks,
                     expand=c(0.01,0))+
  theme_bw()+ # keep minor grid lines
  xlab('AUC')+
  ylab('Residual (observed - expected)')+
  coord_cartesian(xlim=c(lower_auc_plot, NA))+ # focus on near 0.5 and above
  theme(legend.position = 'none')
residual
```


##### Combined plot

```{r combined, fig.height = 8, fig.width = 7}
rounded = rounded + ggtitle('Histogram') +
  xlab('')
#  theme(margin(t=0, r=0, b=0, l=0, unit='pt')) # Not yet checked
residual = residual + ggtitle('Residual')
grid.arrange(rounded, residual, nrow = 2, heights=c(1,0.8))
# export
jpeg('figures/histogram_residuals.jpg', width=5, height=8, units='in', res=500)
grid.arrange(rounded, residual, nrow = 2, heights=c(1,0.8))
invisible(dev.off())
```

## Largest per abstract

The histogram below is restricted to the largest AUC statistic per abstract. We only use the "mean" and "pair" types.

```{r largest}
# largest per abstract
largest = group_by(results, pmid) %>%
  filter(digits > 1, # avoid spikes due to rounding,
         !(type %in% c('Difference','Lower','Upper'))) %>%
  arrange(pmid, desc(auc)) %>%
  slice(1) %>% # largest per abstract
  ungroup()

### rounded histogram
for_histo_largest = mutate(largest, binned = my_bin(auc, digits = 2)) %>% # from lower limit up to upper limit: [lower,upper)
  group_by(binned) %>%
  tally() %>%
  ungroup() %>%
  mutate(fill = ifelse(binned %in% colour, 1, 2))

# smooth Poisson model
model = glm(n ~ ns(binned, df = df), data = for_histo_largest, family=poisson())
f = fitted(model)
for_histo_largest = cbind(for_histo_largest, f) %>%
  mutate(residual = n - f,
         group = 1) %>% # to avoid separate dotted lines by fill
  arrange(binned)

# plot
rounded_largest = ggplot(data = for_histo_largest, aes(x=binned, y=n, fill=factor(fill)))+
  geom_bar(stat='identity', col='grey22', size=0.1)+ # size for thinner lines around bars
  geom_line(data = for_histo_largest, aes(x=binned, y=f, group = group), lty=2)+ # smoothed
  scale_fill_manual(NULL, values=c('dark red','pink','red'))+
  scale_x_continuous(breaks = x.breaks + (0.01)/2, # nudge breaks to top of bar
                     labels = x.breaks,
                     expand=c(0.01,0))+
  theme_bw()+ # keep minor grid lines
  scale_y_continuous(expand=c(0, 150))+ # reduce space to x-axis
  xlab('AUC')+
  ylab('Frequency')+
  coord_cartesian(xlim=c(lower_auc_plot, NA))+ # focus on near 0.5 and above
  theme(legend.position = 'none')
rounded_largest
# export
jpeg('figures/histogram_largest.jpg', width=5, height=4, units='in', res=500)
print(rounded_largest)
invisible(dev.off())
```

##### Combined plot (largest per abstract)

```{r, include=FALSE}
##
residual_largest = ggplot(data = for_histo_largest, aes(x = binned, y = residual, fill = factor(fill)))+
  geom_bar(stat='identity', col='grey22', size=0.1)+
  scale_fill_manual(NULL, values=c('dark red','pink','red'))+
  scale_x_continuous(breaks = x.breaks + (0.01)/2, # nudge breaks to top of bar
                     labels = x.breaks,
                     expand=c(0.01,0))+
  theme_bw()+ # keep minor grid lines
  xlab('AUC')+
  ylab('Residual (observed - expected)')+
  coord_cartesian(xlim=c(lower_auc_plot, NA))+ # focus on near 0.5 and above
  theme(legend.position = 'none')
residual_largest
```


```{r combined_largest, fig.height = 8, fig.width = 7}
##
rounded_largest = rounded_largest + ggtitle('Histogram') +
  xlab('')
residual_largest = residual_largest + ggtitle('Residual')
grid.arrange(rounded_largest, residual_largest, nrow = 2, heights=c(1,0.8))
# export
jpeg('figures/histogram_residuals_largest.jpg', width=5, height=8, units='in', res=500)
grid.arrange(rounded_largest, residual_largest, nrow = 2, heights=c(1,0.8))
invisible(dev.off())
```

## Distribution of AUC statistics for lower and upper confidence limits

```{r intervals, include=FALSE}
### create the data for the rounded histogram
for_histo_ci = filter(results, digits > 1, # avoid spikes due to rounding; include all 1's, even those with 0 decimal places
              (type %in% c('Lower','Upper'))) %>%
  mutate(binned = my_bin(auc, digits = 2)) %>% # from lower limit up to upper limit: [lower,upper)
  group_by(binned, type) %>%
  tally() %>%
  ungroup() %>%
  mutate(fill = ifelse(binned %in% colour, 1, 2))

# smooth Poisson model, one for each limit
df = 4
for_histo_lower = filter(for_histo_ci, type=='Lower')
for_histo_upper = filter(for_histo_ci, type=='Upper')
model_lower = glm(n ~ ns(binned, df = df), data = for_histo_lower, family=poisson())
model_upper = glm(n ~ ns(binned, df = df), data = for_histo_upper, family=poisson())
#
f_lower = fitted(model_lower)
f_upper = fitted(model_upper)
for_histo_lower = cbind(for_histo_lower, f_lower) %>%
  rename('f' = 'f_lower')
for_histo_upper = cbind(for_histo_upper, f_upper)%>%
  rename('f' = 'f_upper')
for_histo_ci = bind_rows(for_histo_lower, for_histo_upper) %>%
  mutate(residual = n - f,
         group = 1) %>% # to avoid separate dotted lines by fill
  arrange(binned)

# histogram
rounded_intervals = ggplot(data = for_histo_ci, aes(x = binned, y = n, fill = factor(fill)))+
  geom_bar(stat='identity', col='grey22', size=0.1)+ # size for thinner lines around bars
  geom_line(data = for_histo_ci, aes(x=binned, y=f, group = group), lty=2)+ # smoothed
  scale_fill_manual(NULL, values=c('dark red','pink','red'))+
  scale_x_continuous(breaks = x.breaks + (0.01)/2, # nudge breaks to top of bar
                     labels = x.breaks,
                     expand=c(0.01,0))+
  theme_bw()+ # keep minor grid lines
  xlab('AUC')+
  ylab('Frequency')+
  scale_y_continuous(expand=c(0, 150))+ # reduce space to x-axis
  coord_cartesian(xlim=c(lower_auc_plot, NA))+ # focus on near 0.5 and above
  theme(legend.position = 'none')+
  facet_wrap(~type, scales='free_y')

## residuals
residual_intervals = ggplot(data = for_histo_ci, aes(x = binned, y = residual, fill = factor(fill)))+
  geom_bar(stat='identity', col='grey22', size=0.1)+
  scale_fill_manual(NULL, values=c('dark red','pink','red'))+
  scale_x_continuous(breaks = x.breaks + (0.01)/2, # nudge breaks to top of bar
                     labels = x.breaks,
                     expand=c(0.01,0))+
  theme_bw()+ # keep minor grid lines
  xlab('AUC')+
  ylab('Residual (observed - expected)')+
  coord_cartesian(xlim=c(lower_auc_plot, NA))+ # focus on near 0.5 and above
  theme(legend.position = 'none')+
  facet_wrap(~type)
```

```{r combined_intervals, fig.height = 8, fig.width = 7}
## combined plot
rounded_intervals = rounded_intervals + ggtitle('Histogram') +
  xlab('')
#  theme(margin(t=0, r=0, b=0, l=0, unit='pt')) # Not yet checked
residual_intervals = residual_intervals + ggtitle('Residual')
grid.arrange(rounded_intervals, residual_intervals, nrow = 2, heights=c(1,0.8))
# export
jpeg('figures/histogram_residuals_intervals.jpg', width=6, height=8, units='in', res=500)
grid.arrange(rounded_intervals, residual_intervals, nrow = 2, heights=c(1,0.8))
invisible(dev.off())
```


## Cumulative plot

The advantage of a cumulative plot is that we do not need to select bins.

```{r cumulative}
# split into facets
gap = 0.02 # gap either side of breaks
breaks = c(0.7, 0.8, 0.9, 1)
labels = paste('(', breaks-gap, ', ', pmin(breaks+gap), ']', sep='')
to_plot = filter(results, str_detect(type, 'Mean|Pair')) %>%
  mutate(facet = case_when(
    (auc>(0.7-gap) & auc <= (0.7+gap)) ~ 1,
    (auc>(0.8-gap) & auc <= (0.8+gap)) ~ 2,
    (auc>(0.9-gap) & auc <= (0.9+gap)) ~ 3,
        auc > (1-gap) ~ 4)
    ) %>%
  filter(!is.na(facet)) %>%
  mutate(facet = factor(facet, levels = 1:4, labels = labels))

## create null distribution with equal steps at round
# get frequency at rounded or not
null = c(breaks, breaks-gap, breaks+gap)
null = null[null<=1]
null_stats = mutate(to_plot,
                   rounded = auc %in% null) %>%
  group_by(facet, rounded) %>%
  summarise(n=n()) %>%
  ungroup()
# spread values according to numbers
null_cdf = NULL
for (k in 1:nrow(null_stats)){
  lower = as.numeric(str_remove(str_split(null_stats$facet[k], pattern=',')[[1]][1], '\\('))
  upper = as.numeric(str_remove(str_split(null_stats$facet[k], pattern=',')[[1]][2], '\\]'))
  if(null_stats$rounded[k]==TRUE){
    numbers = data.frame(aucs = seq(lower, upper, 0.01)) 
    nrow = nrow(numbers)
    numbers = slice(numbers, rep(1:n(), each = round(null_stats$n[k] / nrow)))
    aucs = pull(numbers, aucs)
  }
  if(null_stats$rounded[k]==FALSE){
    aucs = seq(lower+0.0001, upper, length.out = null_stats$n[k])
  }
  frame = data.frame(facet = null_stats$facet[k], aucs = aucs)
     null_cdf = bind_rows(null_cdf, frame)                
}

#
cplot = ggplot(data = to_plot, aes(x = auc)) +
#  stat_ecdf(data = null_cdf, aes(x=auc), geom = "step", size=1, col='grey77') + # not yet working
  stat_ecdf(geom = "step", size=1.05) +
  ylab('Cumulative proportion')+
  xlab('AUC')+
  scale_x_continuous(breaks = seq(0,1,0.01), expand=c(0.01,0))+
  g.theme+
  facet_wrap(~facet, scale='free')
cplot
jpeg('figures/cumulative.jpg', width=5, height=4, units='in', res=500)
print(cplot)
invisible(dev.off())
```

The steps in the plot are due to many results being rounded to 2 decimal places.

This plot combines the "mean" and "pair" types.

## Detailed look at 0.54 to 0.58

```{r}
### create the data for the rounded histogram
colour.narrow = seq(0.54, 0.58, by = 0.01)
x.breaks.narrow = colour.narrow
for_histo_detail = filter(results, 
                   auc>=0.54, auc<=0.58, # narrow range
              !(type %in% c('Difference','Lower','Upper'))) %>%
  mutate(binned = my_bin(auc, digits = 3)) %>% # from lower limit up to upper limit: [lower,upper)
  group_by(binned) %>%
  tally() %>%
  ungroup() %>%
  mutate(fill = ifelse(binned %in% colour, 1, 2)) # not fully working, need fuzzier matching

# histogram
rounded = ggplot(data = for_histo_detail, aes(x = binned, y = n, fill = factor(fill)))+
  geom_bar(stat='identity', col='grey22', size=0.1)+ # size for thinner lines around bars
  scale_fill_manual(NULL, values=c('dark red','pink','red'))+
  scale_x_continuous(breaks = x.breaks.narrow + (0.001)/2, # nudge breaks to top of bar
                     labels = x.breaks.narrow,
                     expand=c(0.01,0))+
  theme_bw()+ # keep minor grid lines
  xlab('AUC')+
  ylab('Frequency')+
  theme(legend.position = 'none')
rounded
```

## Check of 0.55, 0.57 and 1 over time

```{r}
to_plot = filter(results, type %in% c('Mean','Pair')) %>%
  mutate(year = as.numeric(format(date, '%Y'))) %>%
  group_by(year) %>%
  summarise(r1 = sum(auc==0.55),
            r2 = sum(auc==0.57),
            r3 = sum(auc==1),
            n = n()) %>%
  mutate(p1 = r1/n,
            p2 = r2/n,
            p3 = r3/n) %>%
  ungroup() %>%
  filter(n > 100) %>% # at least 100 results
  select('year', starts_with('p')) %>%
  pivot_longer(cols=-year) %>%
  mutate(auc = case_when(
    name == 'p1' ~ '0.55',
    name == 'p2' ~ '0.57',
    name == 'p3' ~ '1'
  ))
#
tplot = ggplot(data = to_plot, aes(x = year, y=value, group = auc, col=auc)) +
  geom_line(size=1.05) +
  ylab('Proportion')+
  xlab('Year')+
  g.theme
tplot
```

## Distributions of AUC statistics by source

Here we check if there is a difference in the distributions according to how the AUC statistic was sourced, e.g., as a confidence interval compared with a "nearby" statistic.

```{r}
to_plot = filter(results, type %in% c("Mean",'Pair')) %>%
  mutate(tauc = 1/(auc+1)) # transform
bplot = ggplot(data = to_plot, aes(x=factor(source), y=auc))+
  geom_boxplot()+
  g.theme+
  xlab('')+
  ylab('AUC')+
  coord_flip()
bplot
```

## Numbers of other AUC statistics in each band

```{r}
counts = NULL
for (bin in unique(for_histo$binned)){
  just_pmid = filter(uncounted, binned == bin) %>%
    select(pmid) %>%
    unique()
  all_for_pmid = left_join(just_pmid, uncounted, by='pmid') %>%
    filter(binned != bin) %>%
    group_by(pmid) %>%
    summarise(n = n(), mean = mean(binned)) %>% # by PMID
    ungroup() %>%
    summarise(meann = mean(n), meana = mean(mean)) %>% # over average per bin
    mutate(binned = bin)
  counts = bind_rows(counts, all_for_pmid)
}
counts = filter(counts, binned > 0.47)

# plot of mean numbers
number_line = ggplot(data = counts, aes(x = binned, y = meann))+
  geom_line(size=1.05, col='goldenrod1')+
  scale_x_continuous(breaks = x.breaks,
                     labels = x.breaks,
                     expand=c(0.01,0))+
  theme_bw()+ # keep minor grid lines
  xlab('AUC')+
  ylab('Mean number of other AUC statistics per abstract')+
  coord_cartesian(xlim=c(lower_auc_plot, NA)) + # focus on near 0.5 and above
  g.theme
number_line
# export
jpeg('figures/numbers_others.jpg', width=5, height=4, units='in', res=500)
print(number_line)
invisible(dev.off())
```

The plot above shows the number of other AUC statistics per abstract for each AUC bin. Abstracts that include a lower AUC statistic tend to have more AUC statistics on average.

```{r}
# line plot of means
mean_binned = ggplot(data = counts, aes(x = binned, y = meana))+
  geom_line(size=1.05, col='dodgerblue')+
  scale_x_continuous(breaks = x.breaks,
                     labels = x.breaks,
                     expand=c(0.01,0))+
  theme_bw()+ # keep minor grid lines
 # scale_y_continuous(expand=c(0, 150))+ # reduce space to x-axis
  xlab('AUC')+
  ylab('Average of other AUC statistics')+
  coord_cartesian(xlim=c(lower_auc_plot, NA)) + # focus on near 0.5 and above
  g.theme
mean_binned
```

The other AUC statistics in an abstract tended to be larger when there was a larger AUC statistic included. This is probably because similar models tended to be compared.
