---
title: "Plots for AUC statistics: sensitivity"
author: "Adrian Barnett"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, comment='', dpi=400)
library(dplyr)
library(tidyr)
library(stringr)
library(splines) # for smooth spline
# graphics
library(ggplot2)
library(gridExtra)
g.theme = theme_bw() + theme(panel.grid.minor = element_blank())
type_colours = c("burlywood2", "dodgerblue", "goldenrod4", "firebrick2")
source('99_functions.R')

# get the data
load('data/analysis_ready_sensitivity.RData') # from 4_combine_results.R

# remove AUC of exactly 1 because of accuracy issues; go to 0.99 because last interval is (0.99, 1]
results = filter(results, auc <= 0.99)

# breaks for histograms
breaks = seq(0, 1.01, 0.01) # up to 1.01 to include 1
# which bars to colour
colour = c(0.7, 0.8, 0.9, 1)
# x-axis
x.breaks = c(0, seq(0.5,1,0.1))
minor_breaks = seq(0.05,0.95,0.1)
# lower limit for plots
lower_auc_plot = 0.48
```

## Histogram

The plot below shows a histogram of all AUC results that were a "mean" or "pair" (so the confidence interval limits are not shown). We restricted the results to those with two or more digits after the decimal place to avoid spikes appearing due to rounding.

The results are binned using a (lower, upper] limit. For example, (0.80, 0.81] is $>$ 0.80 and $\leq$ 0.81.

```{r}
### create the data for the rounded histogram
for_histo = filter(results, 
              (digits > 1 | auc==1), # avoid spikes due to rounding; include all 1's, even those with 0 decimal places
              !(type %in% c('Difference','Lower','Upper'))) %>%
  mutate(binned = my_bin(auc, digits = 2)) %>% # (lower,upper]
  group_by(binned) %>%
  tally() %>%
  ungroup() %>%
  mutate(fill = ifelse(binned %in% colour, 1, 2))

# smooth Poisson model
df = 4
model = glm(n ~ ns(binned, df = df), data = for_histo, family=poisson())
f = fitted(model)
for_histo = cbind(for_histo, f) %>%
  mutate(residual = n - f,
         group = 1) %>% # to avoid separate dotted lines by fill
  arrange(binned)

# histogram
rounded = ggplot(data = for_histo, aes(x = binned, y = n, fill = factor(fill), group = group))+
  geom_bar(stat='identity', col='grey22', size=0.1)+ # size for thinner lines around bars
  geom_line(data = for_histo, aes(x=binned, y=f, group = group), lty=2)+ # smoothed
  scale_fill_manual(NULL, values=c('dark red','pink','red'))+
  scale_x_continuous(breaks = x.breaks + (0.01)/2, # nudge breaks to top of bar
                     labels = x.breaks,
                     expand=c(0.01,0))+
  theme_bw()+ # keep minor grid lines
  scale_y_continuous(expand=c(0, 150))+ # reduce space to x-axis
  xlab('AUC')+
  ylab('Frequency')+
  coord_cartesian(xlim=c(lower_auc_plot, NA))+ # focus on near 0.5 and above
  theme(legend.position = 'none')
rounded
# export
jpeg('figures/histogram_sensitivity.jpg', width=5, height=4, units='in', res=500)
print(rounded)
invisible(dev.off())
```

The dotted line shows a smooth natural spline with `r df` degrees of freedom fitted to the counts in each bin.

#### Residual plot

The plot below is the residuals from the smooth fit to the distribution.

```{r}
residual = ggplot(data = for_histo, aes(x = binned, y = residual, fill = factor(fill)))+
  geom_bar(stat='identity', col='grey22', size=0.1)+
  scale_fill_manual(NULL, values=c('dark red','pink','red'))+
  scale_x_continuous(breaks = x.breaks + (0.01)/2, # nudge breaks to top of bar
                     labels = x.breaks,
                     expand=c(0.01,0))+
  theme_bw()+ # keep minor grid lines
  xlab('AUC')+
  ylab('Residual (observed - expected)')+
  coord_cartesian(xlim=c(lower_auc_plot, NA))+ # focus on near 0.5 and above
  theme(legend.position = 'none')
residual
```


##### Combined plot

```{r combined, fig.height = 8, fig.width = 7}
rounded = rounded + ggtitle('Histogram') +
  xlab('')
#  theme(margin(t=0, r=0, b=0, l=0, unit='pt')) # Not yet checked
residual = residual + ggtitle('Residual')
grid.arrange(rounded, residual, nrow = 2, heights=c(1,0.8))
# export
jpeg('figures/histogram_residuals_sensitivity.jpg', width=5, height=8, units='in', res=500)
grid.arrange(rounded, residual, nrow = 2, heights=c(1,0.8))
invisible(dev.off())
```

